---
title: "Feature Engineering"
author: "Ali Zaidi"
date: "2017/06/05"
output: 
  html_notebook:
    toc: true
    toc_depth: 2
    toc_float: true
editor_options: 
  chunk_output_type: inline
---

## Exploratory Data Analysis and Feature Engineering

Import your favorite libraries and set up your favorite plotting theme.

```{r import-libs, message = F}
library(readr)    # for reading data
library(dplyr)    # for data manipluation
library(dplyrXdf) # for RxXdfData dplyr methods
library(tidyr)    # for data aggregation and spreading
library(purrr)    # for functionals and closures 
library(ggplot2)  # for visualizations
library(maps)     # for maps data
library(scales)   # for units
library(magrittr) # for pipes
library(foreach)  # for parallelization
library(doRSR)    # register backends
theme_set(theme_minimal())      # setting up your default theme

```

Now let's import some data. Our [data source](http://www.dcc.fc.up.pt/~ltorgo/Regression/cal_housing.html) is 1990-census level of housing in California.

We'll use easy the `readr` package to load in our data. You could equivalently use `read.csv`, or `data.table::fread` if you wanted incredible speed. We'll later see how to use data sources using the `RevoScaleR` readers.

Since we're using `readr`, our data is a `tbl`. This means we will get some `dplyr` and `tibble` features, and it'll behave a little differently than a traditional `data.frame`.

```{r import-data, message = F}
housing <- read_csv("~/data/housing.csv")
class(housing)
glimpse(housing)
housing
```

## Visualize Densities

Suppose we want to visualize the distribution of the numeric columns, such as **housing_median_age**. How would you visualize that density?

```{r gather, message = FALSE}
housing %>% keep(is_double) %>% 
  gather %>% 
  ggplot(aes(x = value)) + geom_histogram() + facet_wrap(~key, scales = "free")
```

## Spatial Visualizations

The histograms of the longitude and latitude columns seem like they are in some reasonable range of data. Let's visualize the locations.

```{r spatial-visuals}

housing %>% ggplot(aes(x = longitude, y = latitude)) + 
  geom_point()

```

Looks indeed like California!

We can improve this chart in many ways. Let's first add some transparency so we can get a better sense of the _density_ of points:

```{r alpha}


housing %>% ggplot(aes(x = longitude, y = latitude)) + 
  geom_point(alpha = 0.15)

```

Nice, we can already see some dense clusters for higher population regions like the Bay Area, Sacramento and Los Angeles.

Let's add some additional attributes, like the population:

```{r population}

housing %>% ggplot(aes(x = longitude, y = latitude,
                       size = population)) + 
  geom_point(alpha = 0.15)
```

And now let's see if we can add a gradient fill for another numeric value, like the median housing price:

```{r gradient-fill}

housing %>% ggplot(aes(x = longitude, y = latitude,
                       size = population,
                       colour = median_house_value)) + 
  geom_point(alpha = 0.095)

```

Interesting, we can definitely see the higher price ranges in LA and the Bay Area. Let's see if we can change the colour scheme to get an even better visualization:


```{r colour}

housing %>% ggplot(aes(x = longitude, y = latitude,
                       size = population,
                       colour = median_house_value)) + 
  geom_point(alpha = 0.095) +
  scale_colour_gradient(low = "yellow", high = "red")

```

If you want to plot the points on top of an actual polygon of the state map, you can do that using `geom_map` and using the `map_data` function in ggplot2. I find it a bit messy, but it might be worthwile if you are plotting various regions/states.

While we're at it, let's put some final themes on our plot to make it more aesthetically pleasing.

```{r}

housing %>% 
  mutate(state = "california") %>% 
  ggplot(aes(x = longitude, y = latitude,
             size = population,
             colour = median_house_value)) +
  geom_map(map = filter(map_data("state"), region == "california"),
           aes(map_id = state), 
           fill = "lightgrey",
           colour = "lightgrey") +
  geom_point(alpha = 0.1) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) + 
  scale_size_continuous(labels = scales::comma) +
  scale_colour_gradient(low = "yellow", high = "red",
                        labels = scales::dollar)
```
